MSE and dynamic state variable model with sigma function

### 2018-05-03 doc/MSE_MAR_18.tex
continued work on methods

Indivdual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight. The stock weights are thus calculated from the population numbers as

\begin{equation}
W_i(a,y,s,p) =  \alpha * (N_i (a, y, S, p) * L_{\infty_i} * (1-\exp^{(-K * a)}))^{\beta},
\end{equation}
 and the same is done for the catch weights.

The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model. The Dynamic State Variable model inputs consist of the expected individual catch rates, which are a random variable. This, random variable is normally distributed, and its mean $\hat c_i (a, y, s, p)$  is a function of population size, age dependent catchability $q_i (a)$, and fishing effort $E (y,s,p)$ in any year, season, and area.

\begin{equation}
\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a) * E(y,s,p).
\end{equation}

\subsection{The dynamic-state variable model}
\label{sec2.2}

@@

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}. This was done for 700 individual vessels.

\begin{table}[!h]
\centering
\caption{Model and fishery conditioning.}
\label{t:fishery}
\resizebox{0.6\textwidth}{!}{%
\begin{tabular}{@{}lc@{}} 
\hline
\textbf{model}                              &               \\ \hline
&  \\%& & \\
number of vessels                           &700            \\ 
$\sigma $                                   & 40           \\
number of areas                             & 2            \\
number of seasons                           & 6            \\
intial quota $Q_{sp1}$ and $Q_{sp2}$        & 200 \\%& & \\
&  \\
\textbf{Fishery variable costs}             &               \\ \hline   
&  \\%& & \\
fuel costs (Euro day\textsuperscript{-1})       & 1240  \\ 
gear maintenance (Euro day\textsuperscript{-1}) & 0     \\ 
crew share                                      & 35\% \\ 
landing costs (Euro t\textsuperscript{-1})      & 0    \\ 
&                 \\  
\textbf{Fishery }&                 \\ \hline   
&  \\%& & \\
q                 & 0.0005 \\ 
effort (p,s)      & 1  \\ 
CPUE & mean($pop*q*wts(yy-2:yy)$)\\ 
CPUE dev& $0.08*CPUE$\\ 
&                 \\  \hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependent pricing}
\label{sec2.2.1}
Prices of fish have been assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}. Therefore, the resulting data of mean weights and average prices were used for further analysis. We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively. Thus, following \cite{Zimmermann2011}, the statistical models took the form:  
\begin{equation}
 pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
\end{equation}
where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes. $\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$. We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
 %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models. Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results. All models were fitted independently to each stock.
\begin{table}[!h]
\centering
\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}lcccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
                         & 1     & 2     & 3    & 4   \\ \hline
                      \\                      
species 1 and 2            &  778  & 1354   & 1780  & 2096  \\
\\
\hline 
\end{tabular}
}
\end{table}

@@

>>>>

Individual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight. The weights for individuals in the stock and in the catches are thus calculated as

\begin{equation}
w_i(a,s) =  \alpha * ( L_{\infty_i} * (1-\exp^{(-K * (a+(s/S))}))^{\beta}.
\end{equation}

The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model. The Dynamic State Variable model inputs consist of the expected individual catch rates, which are random variables. These random variables were normally distributed, with means $\hat c_i (a, y, s, p)$ being a function of population size, age dependent catchability $q_i (a)$ in any year, season, and area,

\begin{equation}
\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a)  * w_i(a,s).
\end{equation}
The standard deviations $\Sigma_i (a, y, s, p)$ of the catch distributions were constant fraction  their means, using a ratio $\eta$.

\subsection{fleet dynamics}
\label{sec2.2}

@@

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}. For each year, these forward Monte Carlo Simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area

\begin{table}[!h]
\centering
\caption{Model parameters.}
\label{t:fishery}
\resizebox{0.6\textwidth}{!}{%
\begin{tabular}{@{}lc@{}} 
\hline

\textbf{Population dynamics}                                    &                \\ \hline
&  \\%& & \\
number of vessels                                                      &700            \\ 
$\sigma $                                                                    & 40            \\
number of areas                                                         & 2             \\
number of seasons                                                     & 6             \\
intial quota $Q_{sp1}$ and $Q_{sp2}$                     & 200 \\%& & \\
&  \\

\textbf{Fleet dynamics}                                              &               \\ \hline   
&  \\%& & \\
Number of vessels                                                     & 700 \\ 
fuel costs (Euro day\textsuperscript{-1})                 & 1240    \\ 
gear maintenance (Euro day\textsuperscript{-1})    & 0          \\ 
crew share                                                                 & 35\%    \\ 
landing costs (Euro t\textsuperscript{-1})                & 0           \\ 
&                 \\

\textbf{Fishery }                                                       &          \\ \hline   
&  \\%& & \\
q                                                                                & 0.0005 \\ 
effort (p,s)                                                                & 1  \\ 
Ratio of standard deviations to means $\eta$         & $0.08$\\ 
&                 \\  \hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependent pricing}
\label{sec2.2.1}
Prices of fish were assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}. Following \cite{Zimmermann2011}, therelationship between fish price and body weight was modelled as:  
\begin{equation}
 p_i (a,s) = \bar{ p}_i + \gamma_i \times \frac{w_i (a,s) -\bar{w_i}}{\bar{w_i}},
\end{equation}
where $p_i (a,s) $ is the price of species $i$ at age $a$ and season $s$. The price was a linear function of the weight of inidividuals of species $i$ at age $a$ and season $s$. $\bar{p}_i$ is the price of the species at the mean weight over the age range. The mean weight over the age range is $\bar{w_i}$. $\gamma_i$ gave the price increase when individial mass was increased by $\bar{w_i}$.

\begin{table}[!h]
\centering
\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}lcccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
                         & 1     & 2     & 3    & 4   \\ \hline
                      \\                      
species 1 and 2            &  778  & 1354   & 1780  & 2096  \\
\\
\hline 
\end{tabular}
}
\end{table}

@@

We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
